Aircraft Wildlife collisions

### 2021-07-13 aircraft_wildlife_doc.Rmd
Added birds seen analysis and restructured into analysis of birds strikes, birds seen and effect variable.

@@

# Amount of na for each column
missing <- sapply(birds, function(x) sum(is.na(x)))

barplot(missing)
```

# Analysis of strikes

@@

```{r echo=FALSE, fig.asp=1}
birds.filtered<-birds[!is.na(birds$effect) & !is.na(birds$sky),]
birds.filtered$effectYes<-as.factor(birds.filtered$effect != "None")

prop.table(table(birds.filtered$sky,birds.filtered$effectYes),1)
mosaicplot(prop.table(table(birds.filtered$sky,birds.filtered$effectYes),1))

prop.table(table(birds.filtered$sky,birds.filtered$birds_struck),1)

results show, that clouds do not have an impact on effect variable, 
the proportion of effect=true is actually a little higher for some clouds and overcast

@@

>>>>

#birds seen: remove one wrong entry
which(birds$birds_seen==":-10")
birds$birds_seen[9412] <- NA

factor_seen<-c("2-10", "11-100")
birds$birds_seen <- factor(birds$birds_seen, levels=factor_seen) 
  # drop unused level ":-10" by assigning levels again
levels(birds$birds_seen)

@@

#TODO: get ac mass and num_engs from aircraft model

# Amount of na for each column
missing <- sapply(birds, function(x) sum(is.na(x)))
missing_df <- data.frame("column" = names(missing), 
                         "count" = missing, row.names = NULL)
names(missing_df)[2] <- "count"

ggplot(missing_df) + 
  aes(x=reorder(column, -count), y=count) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Analysis of strikes (birds_struck)

@@

```{r echo=FALSE, fig.asp=1}
#relationship of amount of birds struck with clouds?
birds$effectYes<- as.factor(birds$effect != "None")
birds.filtered<-birds[!is.na(birds$effect) & !is.na(birds$sky),]

prop.table(table(birds.filtered$sky,birds.filtered$birds_struck),1)
ggplot(birds.filtered)+
  aes(x=birds_struck, fill=sky)+
  geom_bar(position = "fill")
```
For 0 birs struck we have highest percentage of no clouds, so clouds might have an affect on avoiding striking birds. Its about 20% higher compared to 1 bird struck.
for birdsstruck by sky we can see that with increasing amount of birds the sky tends to be more more overcast, as no clouds and some clouds decreases

#Clouds with birds seen

there are only 2 levels in the data available for birds seen

```{r echo=FALSE, fig.asp=1}
#relationship of amount of birds seen with clouds?
table(birds$birds_seen)

prop.table(table(birds.filtered$sky,birds.filtered$birds_seen),1)
ggplot(birds.filtered[!is.na(birds.filtered$birds_seen),])+
  aes(x=birds_seen, fill=sky)+
  geom_bar(position = "fill")
```

the proportion of some clouds is nearly the same, but pilots do more often see bigger groups of birds (11-100) when its overcast compared to small groups

```{r echo=FALSE, fig.asp=1}
#relationship of amount of birds seen with daytime?
table(birds$birds_seen)

prop.table(table(birds.filtered$time_of_day,birds.filtered$birds_seen),1)
ggplot(birds.filtered[!is.na(birds.filtered$birds_seen) & !is.na(birds.filtered$time_of_day),])+
  aes(x=birds_seen, fill=time_of_day)+
  geom_bar(position = "fill")
```
## Birds struck out of birds seen

```{r echo=FALSE, fig.asp=1}
#Create new column
#calculate only when seen is not na
df <- birds[!is.na(birds$birds_seen) & !is.na(birds$birds_struck),]
```

check speed

```{r echo=FALSE, fig.asp=1}
# does speed has an influence on seen? the faster the less likely to see ?

spineplot(birds$speed, as.factor(is.na(birds$birds_seen)))

```
The higher the speed, the less likely to see a bird

@@

```{r echo=FALSE, fig.asp=1}
prop.table(table(birds.filtered$sky,birds.filtered$effectYes),1)
ggplot(birds[!is.na(birds$effect) & !is.na(birds$sky),])+
  aes(x=sky, fill=effectYes)+
  geom_bar(position = "fill")
```

results show, that clouds do not have an impact on effect variable, 
the proportion of effect=true is a little higher for some clouds and overcast

check influence of birds seen on effect variable, we assume NA values as if no birds have been seen before the strike occured

```{r echo=FALSE, fig.asp=1}
#anteil der effect=true vergleichen when seen and not seen
#...
effectYes_seen <- ggplot(birds[!is.na(birds$effectYes),]) +
  aes(x=is.na(birds_seen), fill=effectYes) +
  geom_bar(position="fill")

#differences in kind of effect?
effect_seen <- ggplot(birds[!is.na(birds$effect) & birds$effect != "None" ,]) +
  aes(x=is.na(birds_seen), fill=effect) +
  geom_bar(position="fill")

effectYes_seen + effect_seen

proportion of effect is a little higher when birds were seen
